Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-06-06 Multiresolution_framework/ltx/manuscript.tex
got rid of bspline expansion

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale using the two-scale relation pair defined in Eqs.~(\eqref{eq:MRA-TwoScalepair1}-\eqref{eq:MRA-TwoScalepair2coefs}). For scaling functions this can be done  by using Eq.~\eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
 l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see \ref{ap:Induction}). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using Eq.~\eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Fig.~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
 		\includegraphics[scale=1]{./Graph/fig5.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (dotted line) at $j=0$ is obtained using 625 scaling functions (solid lines) at $j=4$. (\textbf{B}) The wavelet function (dotted line) at $j=0$ is obtained using 1250 scaling functions (solid lines) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}

@@

\newpage
\appendix
\section{Inner product of two B-spline scaling functions}\label{ap:InnerProductOfBsplines}
In this section, an analytic formula is derived for the inner product of two B-spline scaling functions. Consider two B-spline scaling functions of order $m$ and $m'$. The support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e.
\begin{align}\label{eq:APP-SymmetrySplieEq}
 N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:APP-SymmetrySplieEq} is 
\begin{align}
 N_{m}\left(r\right)=N_{m}\left(m-r\right).
\end{align}
Therefore
\begin{align}
\int_{-\infty}^{+\infty}N_{m}\left(r-l_{1}\right)N_{m'}\left(r-l_{2}\right)dr&=\int_{-\infty}^{+\infty}N_{m}\left(m-r+l_{1}\right)N_{m'}\left(r-l_{2}\right)dr \nonumber \\
&=\int_{-\infty}^{+\infty}N_{m}\left(m+l_{1}-l_{2}-u\right)N_{m'}\left(u\right)du \nonumber \\
&=\left(N_m \ast N_{m'}\right) \left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m'+l_{2}-l_{1}\right).
\end{align}
\section{Scaling function expansion}\label{ap:Induction}  
In this appendix we show by induction that how the scaling function at $j=0$ can be written in terms of the scaling function at level $j$. For $n_1$ in  Eq.~\eqref{eq:MRA-ConvertingFormulaForscalingFunctions} we have
\begin{align}
 N_{m}\left(r\right)&=\sum_{n_1=0}^{m} \left[\mathbf p\right]_{n_1} N_{m}\left(2r-n_1\right) \label{eq:App-TwoScalepair1},
  \end{align}
which clearly holds from two-scale relation pair (see Eq.~\eqref{eq:MRA-TwoScalepair1}). Now assuming Eq.~\eqref{eq:MRA-ConvertingFormulaForscalingFunctions} for $n_{j-1}$ we can write 
\begin{align}\label{eq:app-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_{j-1}=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_{j-1}}N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}}),
\end{align}
in which $N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})$ can be written in terms of $N_m(2^{j}r)$ as
\begin{align}\label{eq:AppendixNj-1intermsofNj}
 N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})&=\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr-2\times k_{n_1,n_2, \dots, n_{j-1}}-n_j)\nonumber \\
&=\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr- k_{n_1,n_2, \dots, n_{j}}).
\end{align}
Substituting Eq.~\eqref{eq:AppendixNj-1intermsofNj} into Eq.~\eqref{eq:app-ConvertingFormulaForscalingFunctions} completes the proof.

%\section{Lower bound on likelihood function}\label{ap:Lowerbound}
%A derivation of the lower bound on the likelihood function that is based on \citet{Bishop2006} is provided in this Appendix. We want to maximize the likelihood function, 
%\begin{equation}
 %p(\mathbf Y;\boldsymbol\theta)=\int_{\mathbf X}p(\mathbf X,\mathbf Y;\boldsymbol\theta)\mathrm{d}\mathbf X,
%\end{equation}
%which is difficult to perform directly. Maximizing the likelihood in terms of the parameter set is equivalent to the log-likelihood maximization, as both the likelihood and log-likelihood functions share the same maximizer due to the monotonicity of the logarithm function. For any choice of distribution defined over the hidden variables, $q(\mathbf X)$, the following decomposition holds
%\begin{equation}\label{eq:EM-loglikelihooddecomposition}
 %\ln~p(\mathbf Y;\boldsymbol\theta)=\mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p),
%\end{equation}
%where
%\begin{align}
 %\mathcal{L}(q,\boldsymbol\theta)&=\int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X,\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \label{eq:EM-decomposition1}\\
%\mathrm{KL}(q\parallel p)&=-\int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X. \label{eq:EM-decomposition2}
%\end{align}
%Eq.~\eqref{eq:EM-loglikelihooddecomposition} can be verified by applying the probability product rule to the joint distribution $p(\mathbf X,\mathbf Y;\boldsymbol\theta)$ in Eq.~\eqref{eq:EM-decomposition1} giving
%\begin{align}
 % \mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p)&= \int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)p(\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \nonumber\\
	%&- \int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X\nonumber\\
	%&= \int_{\mathbf X}q(\mathbf X)\ln p(\mathbf Y; \boldsymbol\theta)\mathrm{d}\mathbf X.
%\end{align}
%Noting  $q(\mathbf X)$ is a normalized distribution that integrates to one gives the log-likelihood function. We also note that Eq.~\eqref{eq:EM-decomposition2} is the Kullback-Liebler distance between distributions $p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$ and $q(\mathbf X)$. Therefore,
%\begin{equation}
 % \mathrm{KL}(q\parallel p)\ge 0.
%\end{equation}   
%The equality holds if and only if $q(\mathbf X)=p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$. It follows from  Eq.~\eqref{eq:EM-loglikelihooddecomposition}, that the quantity $\mathcal{L}(q,\boldsymbol\theta)$ is a lower bound on the log-likelihood function, i.e.
%\begin{equation}
 %\mathcal{L}(q,\boldsymbol\theta)\le \ln~p(\mathbf Y\mid\boldsymbol\theta)
	%\end{equation}

>>>>

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale using the two-scale relation pair defined in Eqs.~(\eqref{eq:MRA-TwoScalepair1}-\eqref{eq:MRA-TwoScalepair2coefs}). 
% For scaling functions this can be done  by using Eq.~\eqref{eq:MRA-TwoScalepair1} recursively, giving
% \begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
%  &N_m(r)=\nonumber \\
% &\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
% where 
%  l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
% The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see \ref{ap:Induction}). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using Eq.~\eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Fig.~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
% \begin{figure}[!t] 
%  	\centering
%  		\includegraphics[scale=1]{./Graph/fig5.pdf}
%  		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (dotted line) at $j=0$ is obtained using 625 scaling functions (solid lines) at $j=4$. (\textbf{B}) The wavelet function (dotted line) at $j=0$ is obtained using 1250 scaling functions (solid lines) at $j=4$.}
%  	\label{fig:MRA-BasisDecomposition}
%  \end{figure}  

@@

\newpage
\appendix
\section{Inner product of two B-spline scaling functions}\label{ap:InnerProductOfBsplines}
In this section, an analytic formula is derived for the inner product of two B-spline scaling functions. Consider two B-spline scaling functions of order $m$ and $m'$. The support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e.
\begin{align}\label{eq:APP-SymmetrySplieEq}
 N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right).
A direct consequence of Eq.~\eqref{eq:APP-SymmetrySplieEq} is 
 N_{m}\left(r\right)=N_{m}\left(m-r\right).
Therefore
\int_{-\infty}^{+\infty}N_{m}\left(r-l_{1}\right)N_{m'}\left(r-l_{2}\right)dr&=\int_{-\infty}^{+\infty}N_{m}\left(m-r+l_{1}\right)N_{m'}\left(r-l_{2}\right)dr \nonumber \\
&=\int_{-\infty}^{+\infty}N_{m}\left(m+l_{1}-l_{2}-u\right)N_{m'}\left(u\right)du \nonumber \\
&=\left(N_m \ast N_{m'}\right) \left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m+l_{1}-l_{2}\right) \nonumber \\
&=N_{m+m'}\left(m'+l_{2}-l_{1}\right).
% \section{Scaling function expansion}\label{ap:Induction}  
% In this appendix we show by induction that how the scaling function at $j=0$ can be written in terms of the scaling function at level $j$. For $n_1$ in  Eq.~\eqref{eq:MRA-ConvertingFormulaForscalingFunctions} we have
%  N_{m}\left(r\right)&=\sum_{n_1=0}^{m} \left[\mathbf p\right]_{n_1} N_{m}\left(2r-n_1\right) \label{eq:App-TwoScalepair1},
% which clearly holds from two-scale relation pair (see Eq.~\eqref{eq:MRA-TwoScalepair1}). Now assuming Eq.~\eqref{eq:MRA-ConvertingFormulaForscalingFunctions} for $n_{j-1}$ we can write 
% \begin{align}\label{eq:app-ConvertingFormulaForscalingFunctions}
%  &N_m(r)=\nonumber \\
% &\sum_{n_1,n_2, \dots n_{j-1}=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_{j-1}}N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}}),
% in which $N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})$ can be written in terms of $N_m(2^{j}r)$ as
% \begin{align}\label{eq:AppendixNj-1intermsofNj}
%  N_m(2^{j-1}r-k_{n_1,n_2, \dots, n_{j-1}})&=\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr-2\times k_{n_1,n_2, \dots, n_{j-1}}-n_j)\nonumber \\
% &=\sum_{n_j=0}^m \left[\mathbf p\right]_{n_j}N_m(2^jr- k_{n_1,n_2, \dots, n_{j}}).
% Substituting Eq.~\eqref{eq:AppendixNj-1intermsofNj} into Eq.~\eqref{eq:app-ConvertingFormulaForscalingFunctions} completes the proof. 
%\section{Lower bound on likelihood function}\label{ap:Lowerbound}
%A derivation of the lower bound on the likelihood function that is based on \citet{Bishop2006} is provided in this Appendix. We want to maximize the likelihood function, 
%\begin{equation}
 %p(\mathbf Y;\boldsymbol\theta)=\int_{\mathbf X}p(\mathbf X,\mathbf Y;\boldsymbol\theta)\mathrm{d}\mathbf X,
%\end{equation}
%which is difficult to perform directly. Maximizing the likelihood in terms of the parameter set is equivalent to the log-likelihood maximization, as both the likelihood and log-likelihood functions share the same maximizer due to the monotonicity of the logarithm function. For any choice of distribution defined over the hidden variables, $q(\mathbf X)$, the following decomposition holds
%\begin{equation}\label{eq:EM-loglikelihooddecomposition}
 %\ln~p(\mathbf Y;\boldsymbol\theta)=\mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p),
%\end{equation}
%where
 %\mathcal{L}(q,\boldsymbol\theta)&=\int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X,\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \label{eq:EM-decomposition1}\\
%\mathrm{KL}(q\parallel p)&=-\int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X. \label{eq:EM-decomposition2}
%Eq.~\eqref{eq:EM-loglikelihooddecomposition} can be verified by applying the probability product rule to the joint distribution $p(\mathbf X,\mathbf Y;\boldsymbol\theta)$ in Eq.~\eqref{eq:EM-decomposition1} giving
 % \mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p)&= \int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)p(\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \nonumber\\
	%&- \int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X\nonumber\\
	%&= \int_{\mathbf X}q(\mathbf X)\ln p(\mathbf Y; \boldsymbol\theta)\mathrm{d}\mathbf X.
%Noting  $q(\mathbf X)$ is a normalized distribution that integrates to one gives the log-likelihood function. We also note that Eq.~\eqref{eq:EM-decomposition2} is the Kullback-Liebler distance between distributions $p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$ and $q(\mathbf X)$. Therefore,
%\begin{equation}
 % \mathrm{KL}(q\parallel p)\ge 0.
%\end{equation}   
%The equality holds if and only if $q(\mathbf X)=p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$. It follows from  Eq.~\eqref{eq:EM-loglikelihooddecomposition}, that the quantity $\mathcal{L}(q,\boldsymbol\theta)$ is a lower bound on the log-likelihood function, i.e.
%\begin{equation}
 %\mathcal{L}(q,\boldsymbol\theta)\le \ln~p(\mathbf Y\mid\boldsymbol\theta)
	%\end{equation}
